---
title: "H1b Employer Data Analysis"
author: "Narender Tumu"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r Loading packages,message=FALSE,warning=FALSE,include=FALSE}
# Loading the necessary packages

library(tidyr)
library(knitr)
library(kableExtra)
library(tidyverse)
library(tidymodels)
library(readxl)
library(readr)
library(dplyr)
library(ggplot2)
library(pscl)
library(glmnet)
library(treemap)
library(GGally)

```



## Loading the "H1b_data.RData" file which was obtained after cleaning the raw data files. For the process of cleaning the data navigate to [Data Cleaning](https://github.com/NarenderTumu/H1B-employer-data-analysis/tree/main/Files/Data%20Cleaning)



```{r Loading data,warning=FALSE,include=FALSE}
load("~/STA631/H1B-employer-data-analysis/Files/Data Cleaning/H1b_data.RData")
```

#### Below are the Data dictionaries to understand the data better

**Data Dictionary for the H1B employer data**

```{r creating a Data Dictionary,echo=FALSE}
# Define the data dictionary as a data frame
dict <- data.frame(
  Variable = names(H1b_data),
  Description = c("Petitioner’s five-digit ZIP code ",
                  "This variable indicates the fiscal year in which the H-1B petition was filed. Fiscal year in the United States starts from October 1st and ends on September 30th of the following year.", 
                  "Petitioner’s firm/employer name", 
                  "Number of approvals when the employer filing petition for the first time to the worker",
                  "Number of denails when the employer filing petition for the first time to the worker",
                  "Number of approvals when the employer is filing the petition for the worker who was already having an approved visa for the same employer and wishing to extend the workers time period",
                  "Number of denials when the employer is filing the petition for the worker who was already having an approved visa for the same employer and wishing to extend the workers time period",
                  "North American Industry Classification System Code: A character string that stands for an industry classification within the North American Industry Classification System,For more information on the NAICS, visit the [U.S. Census Bureau’s North American Industry Classification Code webpage](https://www.census.gov/naics/)",
                  "Petitioner’s state",
                  "Petitioner’s City",
                  "Population of the place by Zipcode"), 
Datatype = sapply(H1b_data, class),stringsAsFactors = FALSE)


kable(dict,align = "lccc", row.names = FALSE ) %>%
kable_styling("bordered",full_width = FALSE,position = "center")
```




**The following table provides detailed information on the definitions of NAICS codes**



```{r,echo=FALSE}
NAICS_dictionary<-data.frame(Sector=c("11","21","22","23","31-33","42","44-45","48-49","51","52","53","54","55","56","61","62","71","72","81","92","99"),Definition=c("Agriculture, Forestry, Fishing and Hunting","Mining, Quarrying, and Oil and Gas Extraction","Utilities","Construction","Manufacturing","Wholesale Trade","Retail Trade","Transportation and Warehousing","Information","Finance and Insurance","Real Estate and Rental and Leasing","Professional, Scientific, and Technical Services","Management of Companies and Enterprises","Administrative and Support and Waste Management and Remediation Services","Educational Services","Health Care and Social Assistance","Arts, Entertainment, and Recreation","Accommodation and Food Services","Other Services (except Public Administration)","Public Administration","Unknown"))

kable(NAICS_dictionary, row.names = FALSE ) %>%
  kable_styling("bordered",full_width = FALSE,position = "center")

```




we have variables that are numeric, character and integer. Let's convert some of the variables to use them for analysis


```{r}
# converting the Fiscal_Year, NAICS variables from numeric to factor

H1b_data$Fiscal_Year<-as.factor(H1b_data$Fiscal_Year)

H1b_data$NAICS<-as.factor(H1b_data$NAICS)


# converting the State, City variables from Charatcer to factor

H1b_data$State<-as.factor(H1b_data$State)

H1b_data$City<-as.factor(H1b_data$City)
```



Checking the Summmary statistics of all the numeric variables in the dataset using select() and summary() functions



```{r,echo=FALSE}
H1b_data %>% select(Initial_Approvals,Initial_Denials,Continuing_Approvals,Continuing_Denials,Total_population)%>%
  summary()
```


* Using the table() function for all the factor variables to get the frequency distribution of each level


* We can see that from the below plot, most of the employers belongs to Professional, Scientific, and Technical Services category


```{r,echo=FALSE}
NAICS_frequency<-data.frame(table(H1b_data$NAICS))


ggplot(NAICS_frequency,aes(x=reorder(Var1,Freq),y=Freq))+
  geom_col(color="black",fill="blue")+
  labs(title="Number of employers based on their NAICS classification",
       x="NAICS code",
       y="Number of employers",
       caption = "Data source: USCIS H1B Employer Datahub 2017-2021")+
  theme_light()+
  theme(plot.title = element_text(hjust = 0.5),axis.text.x = element_text(color = "black"),axis.text.y = element_text(color = "black"))

```



* We can see from the below chart, that California has the highest number of employees filing for H1B visa followed by New York, Texas, New Jersey


```{r,echo=FALSE}
State_frequency<-data.frame(table(H1b_data$State))


ggplot(State_frequency,aes(x=reorder(Var1,Freq),y=Freq,fill=Freq))+
  geom_col(color="black")+
  scale_fill_gradient(low = "blue", high = "orange")+
  labs(title="Number of employers filed petitions from each state (2017-2021)",
       x="State",
       y="Number of employers",
       caption = "Data source: USCIS H1B Employer Datahub 2017-2021")+
    theme_light()+
  theme(plot.title = element_text(hjust = 0.5),axis.text.x = element_text(angle = 90, hjust = 1,colour = "black"),axis.text.y = element_text(colour = "black"))
```

* We can see from the below tree map of top 30 cities in US based on the number of employers filing H1B petitions from 2017-2021, where New York, San Francisco, Houston, Chicago, Boston stands in the top

```{r,echo=FALSE,warning=FALSE}

City_frequency<-data.frame(table(H1b_data$City))
City_frequency<-City_frequency%>% arrange(desc(Freq))

Top30_cities<-City_frequency[1:30,]

Top30_cities$Freq<-as.numeric(Top30_cities$Freq)

treemap(Top30_cities,
        index = "Var1",
        vSize = "Freq",
        vColor="Freq",
        type = "value",
        palette = "YlOrBr",
        title = "Top 30 US Cities by employers filing H1B Petitions, 2017-2021")
```


#### Checking the distribution of variables and correlations between the numerical variables

 * Here are the histograms for all the numeric variables. We can observe that the variables Initial_Approvals, Initial_Denials, Continuing_Approvals, and Continuing_Denials are heavily right skewed, while the Total_population variable is only slightly right skewed.


 
```{r,echo=FALSE}
hist(H1b_data$Initial_Approvals)
hist(H1b_data$Initial_Denials)
hist(H1b_data$Continuing_Approvals)
hist(H1b_data$Continuing_Denials)
hist(H1b_data$Total_population)
```


* Below we can see the output of ggpairs() which gives histograms and scatter plots of all the numeric variables with their respective correlation-coefficients from which we can see there is some correlation between Initial_Approvals and Continuing Approvals. 


```{r,warning=FALSE,echo=FALSE}
H1b_data%>%select(Initial_Approvals,Initial_Denials,Continuing_Approvals,Continuing_Denials,Total_population)%>%
  ggpairs(na.rm=TRUE)
```


* This Scatter plot below explains the relationship between Initial and Continuing Approvals

```{r,echo=FALSE}
ggplot(H1b_data,aes(x=Continuing_Approvals,y=Initial_Approvals))+
  geom_point(color="red")+
  labs(title = "Scatterplot between Initial and Continuing Approvals",
       x="Continuing Approvals",
       y="Initial Approvals",
       caption = "Data source: USCIS H1B Employer Datahub 2017-2021")+
  theme(plot.title = element_text(hjust = 0.5),plot.caption.position = "panel")
```

```{r}
#H1b_data$Initial_Approvals<-H1b_data$Initial_Approvals+1
#H1b_data$Continuing_Approvals<-H1b_data$Continuing_Approvals+1
```



```{r}

#model<-lm(log(Initial_Approvals)~log(Continuing_Approvals),data = H1b_data)
#summary(model)

```

```{r}
#plot(model)
```



## Up to this point, I have attempted to construct various statistical models, such as simple and multiple linear regression models, logistic regression, poisson regression, zero-inflated poisson regression, negative binomial regression, and zero-inflated negative binomial regression models. These models were intended to predict the response variable "Initial_Approvals" using a range of predictors. Unfortunately, my attempts to predict the response variable have not been particularly successful. It's possible that I need to change my approach or look for errors in my model building. There is still a great deal of future work that needs to be done on this project. If you have any suggestions or would like to discuss the project further, [please to reach out to me](https://github.com/NarenderTumu/H1B-employer-data-analysis/issues/1).


