---
title: "H1b Employer Data Analysis"
author: "Narender Tumu"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r Loading packages,message=FALSE,warning=FALSE,include=FALSE}
# Loading the necessary packages

library(tidyr)
library(knitr)
library(kableExtra)
library(tidyverse)
library(tidymodels)
library(readxl)
library(readr)
library(dplyr)
library(ggplot2)
library(pscl)
library(glmnet)

```



## Loading the "H1b_data.RData" file which was obtained after cleaning the raw data files. For the process of cleaning the data navigate to [Data Cleaning](https://github.com/NarenderTumu/H1B-employer-data-analysis/tree/main/Files/Data%20Cleaning)



```{r Loading data,warning=FALSE,include=FALSE}
load("~/STA631/H1B-employer-data-analysis/Files/Data Cleaning/H1b_data.RData")
```



**Data Dictionary for the H1B employer data**

```{r creating a Data Dictionary,echo=FALSE}
# Define the data dictionary as a data frame
dict <- data.frame(
  Variable = names(H1b_data),
  Description = c("Petitioner’s five-digit ZIP code ",
                  "This variable indicates the fiscal year in which the H-1B petition was filed. Fiscal year in the United States starts from October 1st and ends on September 30th of the following year.", 
                  "Petitioner’s firm/employer name", 
                  "Number of approvals when the employer filing petition for the first time to the worker",
                  "Number of denails when the employer filing petition for the first time to the worker",
                  "Number of approvals when the employer is filing the petition for the worker who was already having an approved visa for the same employer and wishing to extend the workers time period",
                  "Number of denials when the employer is filing the petition for the worker who was already having an approved visa for the same employer and wishing to extend the workers time period",
                  "North American Industry Classification System Code: A character string that stands for an industry classification within the North American Industry Classification System,For more information on the NAICS, visit the [U.S. Census Bureau’s North American Industry Classification Code webpage](https://www.census.gov/naics/)",
                  "Petitioner’s state",
                  "Petitioner’s City",
                  "Population of the place by Zipcode"), 
Datatype = sapply(H1b_data, class),stringsAsFactors = FALSE)


kable(dict,align = "lccc", row.names = FALSE ) %>%
  kable_styling("bordered", full_width = FALSE)

```




**The following table provides detailed information on the definitions of NAICS codes**



```{r,echo=FALSE}
NAICS_dictionary<-data.frame(Sector=c("11","21","22","23","31-33","42","44-45","48-49","51","52","53","54","55","56","61","62","71","72","81","92","99"),Definition=c("Agriculture, Forestry, Fishing and Hunting","Mining, Quarrying, and Oil and Gas Extraction","Utilities","Construction","Manufacturing","Wholesale Trade","Retail Trade","Transportation and Warehousing","Information","Finance and Insurance","Real Estate and Rental and Leasing","Professional, Scientific, and Technical Services","Management of Companies and Enterprises","Administrative and Support and Waste Management and Remediation Services","Educational Services","Health Care and Social Assistance","Arts, Entertainment, and Recreation","Accommodation and Food Services","Other Services (except Public Administration)","Public Administration","Unknown"))

kable(NAICS_dictionary,align = "lccc", row.names = FALSE ) %>%
  kable_styling("bordered", full_width = FALSE)
```




we have variables that are numeric, character and integer. Let's convert some of the variable to use them for analysis


```{r}
# converting the Fiscal_Year, NAICS variables from numeric to factor

H1b_data$Fiscal_Year<-as.factor(H1b_data$Fiscal_Year)

H1b_data$NAICS<-as.factor(H1b_data$NAICS)


# converting the State, City variables from Charatcer to factor

H1b_data$State<-as.factor(H1b_data$State)

H1b_data$City<-as.factor(H1b_data$City)
```



Checking the Summmary statistics of all the numeric variables in the dataset using select() and summary() functions



```{r,echo=FALSE}
H1b_data %>% select(Initial_Approvals,Initial_Denials,Continuing_Approvals,Continuing_Denials,Total_population)%>%
  summary()
```


### Using the table() function for all the factor variables to get the frequency distribution of each level


### We can see that from the below plot, most of the employers belongs to Professional, Scientific, and Technical Services category


```{r,echo=FALSE}
NAICS_frequency<-data.frame(table(H1b_data$NAICS))


ggplot(NAICS_frequency,aes(x=reorder(Var1,Freq),y=Freq))+
  geom_col(color="black",fill="blue")+
  labs(title="Number of employers based on their NAICS classification",
       x="NAICS code",
       y="Number of employers")+
  theme_light()+
  theme(plot.title = element_text(hjust = 0.5))

```









```{r}
H1b_data$Initial_Approvals<-H1b_data$Initial_Approvals+1
H1b_data$Continuing_Approvals<-H1b_data$Continuing_Approvals+1
```



```{r}

model<-lm(log(Initial_Approvals)~log(Continuing_Approvals),data = H1b_data)
summary(model)

```

```{r}
plot(model)
```






