---
title: "Data Preperation for visualization"
author: "Narender"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### I am trying to generate an interactive US map visualization using R. The visualization will allow users to hover over and select a specific state or city of interest to obtain the numerical count of approvals and denials for employers operating in that area.


* To start with, I have loaded the [data](https://github.com/NarenderTumu/H1B-employer-data-analysis/blob/main/Files/Data%20Cleaning/H1b_data.RData) I am using for this project and also a new data set that was obtained from [simplemaps](https://simplemaps.com/data/us-cities) website which I am going to use to get the latitudes and longitudes data that will be helpful to build a map visualization.



* I have also grouped and summarized the H1b_data, that is I have grouped the data by Employer and State variables and then summarized their approvals and denials variables by using sum()function. A sample of that data can be seen below


```{r,echo=FALSE,warning=FALSE,message=FALSE}
# Loading necessary packages

library(dplyr)
library(kableExtra)
library(readr)
# Loaded the H1b data

load("~/STA631/H1B-employer-data-analysis/Files/Data Cleaning/H1b_data.RData")


# Grouped the data by employer, state, city and summarized the approvals and denials numerics

Grouped_data<-H1b_data%>%group_by(Employer,State,City)%>%summarise(total_initial_approvals=sum(Initial_Approvals),total_initial_denials=sum(Initial_Denials),total_Continuing_approvals=sum(Continuing_Approvals),total_continuing_denials=sum(Continuing_Denials),.groups = "drop")

# subset some observations to display a sample table using kable function

Grouped_subset<-Grouped_data[1:5,]

kable(Grouped_subset)%>%kable_styling("bordered",full_width = FALSE)
```



* Below is the sample of data that I obtained from [simplemaps](https://simplemaps.com/data/us-cities) and also cleaned it based on the requirements for the visualization



```{r,echo=FALSE}

# Downloaded the data from the given website and loaded it in R using read_csv

uscities<-read_csv("~/STA631/H1B-employer-data-analysis/Files/Visualization Dashboard/uscities.csv",show_col_types = FALSE)


# Subset only the required columns for this visualization which is City, State, latitude and longitude

uscities<-uscities[,c(1,4,7,8)]


# rename the city and state variable to match the names in the H1b grouped data

uscities<-uscities%>%rename(City=city,State=state_name)

# Changing the City variable observations to upper case to match the city variable in H1b grouped data

uscities$City<-toupper(uscities$City)


# subset some observations to display a sample table using kable function

uscities_subset<-uscities[1:5,]

kable(uscities_subset)%>%kable_styling("bordered",full_width = FALSE)
```




* Here is the merged and final data that I am going to use to create a US map visualization. To access this data navigate to this [location](https://github.com/NarenderTumu/H1B-employer-data-analysis/tree/main/Files/Visualization%20Dashboard)

```{r,echo=FALSE}

# Merged both data sets using merge function

Grouped_data2<-merge(Grouped_data,uscities,by=c("State","City"),all.x = TRUE)

# Removing missing values from the data

complete_observations<-complete.cases(Grouped_data2)

Data<-Grouped_data2[complete_observations,]

# subset some observations to display a sample table using kable function

Data_subset<-Data[1:5,]

kable(Data_subset)%>%kable_styling("bordered",full_width = FALSE)

save(Data,file = "~/STA631/H1B-employer-data-analysis/Files/Visualization Dashboard/Data.RData")

```









